A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-01-15 BSS7782.tex
worked on text

When adopting a mass of the tertiary star of $m_3 = 1.4$\,\MSun, we
can estimate the initial parameters for the inner binary as well as
the orbit of the outer star after mass transfer. We first calculate
the stellar radius as a function of core mass. In
figure\,\ref{fig:tertiarymass_vs_size} we present this relation
calculated using the {\tt SeBa} stellar evolution code
\cite{1996A&A...309..179P} (dark blue curve).  The interruption in
this curve, aroung a core mass of $m_{\rm core} \sim 0.5$\,\MSun\, is
a result of the evolution along the horizontal branch, where the core
of the star continues to grow but the radius actually shrinks.

Adopting masses for the inner binary $m_1=1.1$\,\MSun\, and
$m_2=0.9$\,\MSun\, we can calculate the outer orbital separation at
the onset of Roche-lobe overflow $a_{\rm out}$, and subsequently the
maximum orbital separation for the inner binary for which the orbit is
stable and a circum binary disk can form. These two limits are
prosented as the light-blue and light green curves in
figure\,\ref{fig:tertiarymass_vs_size}.  The alotted region of
parameter space is then above the dahsed horizontal line and to the
right of the vertical dotted line.

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star.  The change in
orbital separation due to non-conservative mass transfer can be
expressed in terms of the mass of the outer star before and after mass
transfer $m_3$ and $m'_3$, the total mass in the inner binary before
($m_{\rm in}$) and after having accreted ($m'_{\rm in}$) and the
amount of angular momentum lost per unit mass $\eta \simeq 3$
\cite{1995A&A...296..691P}.
\begin{equation}
  {a \over a_0} = \left( {m'_3 m'_{\rm in} \over m_3 m_{\rm in}} \right)^{-2}
  \left( {m'_3 + m'_{\rm in} \over m_3 + m_{\rm in}} \right)^{2\eta + 1}
\end{equation}
In figure\,\ref{fig:tertiarymass_vs_size} we present this relation
with the top brown curve, which subsequently gives an estimate of the
orbital separation of the white dwarf around the twin blue straggler
7782.

>>>>

When adopting a mass of the tertiary star of $m_3 = 1.4$\,\MSun, we
can constrain the initial parameters for the inner binary as well as
the orbit of the outer star after mass transfer. We first calculate
the stellar radius as a function of core mass. In
figure\,\ref{fig:tertiarymass_vs_size} we present this relation
calculated using the {\tt SeBa} stellar evolution code
\cite{1996A&A...309..179P} (dark blue curve).  The interruption in
this curve, around a core mass of $m_{\rm core} \sim 0.5$\,\MSun\, is
a result of the evolution along the horizontal branch, where the core
of the star continues to grow but the radius actually shrinks.
Roche-lobe overflow in this phase is not expected to happen, because
it would already have happened in an earlier evolutionary state of the
donor star, when it was bigger.

Adopting masses for the inner binary $m_1=1.1$\,\MSun\, and
$m_2=0.9$\,\MSun\, we can calculate the outer orbital separation at
the onset of Roche-lobe overflow $a_{\rm out}$, and subsequently the
maximum orbital separation for the inner binary for which the orbit is
stable and a circum-binary disk can form. These two limits are
presented as the light-blue and light green curves in
figure\,\ref{fig:tertiarymass_vs_size}.  The alotted region of
parameter space is then above the dashed horizontal line and to the
right of the vertical dotted line.

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after the phase of
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a'$) and after mass transfer
($a$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a \over a'} = \left( {m'_3 m'_{\rm in} \over m_3 m_{\rm in}} \right)^{-2}
  \left( {m'_3 + m'_{\rm in} \over m_3 + m_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides an
estimate of the orbital separation of the white dwarf around the twin
blue straggler 7782. Tidal effects during the episode of mass tranfer
have probably circularized the orbit, althoug some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small eccentricity of $e \ll 0.1$.
